Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-11-08 manuscript/reviewer_reply.Rmd

In general, I thought this paper was interesting and well written. These results add to the increasing literature supporting the idea that ecological forecasting in time/space is difficult. I think the recommendations/best practice bullet points described by the authors are essential for anyone interested in forecasting this kind of data. All of my points below can be considered minor issues, which I think will improve the paper’s clarity.

Introduction:
Line 41: I think this is generally more common in terrestrial ecology – in fisheries these SDM models with more of a theoretical background don’t seem to be used as much

Line 62: I really like the recommendations / clarity in Box 1, and think these will be accessible to most ecologists

"Line 71-73: I’m re-reading this after commenting on data, and am now a little confused. The bird responses are annual snapshots during the breeding season? And I thought the environmental data you used was annual or quarterly, not monthly?
https://www.mbr-pwrc.usgs.gov/bbs/grass/genintro.htm"

--BBS data are generally viewed as annual data, even though the observations happen on a single day. The presence of an individual species is primarily a function of suitable habitat and, for many species, timing of migration. Environmental variables with a shorter temporal scale could likely improve models, such as incorporating cool weather days into the observation model, and would be worth pursuing in a different study.

@@

Methods:
Line 92: For those not familiar with the BBS protocol, maybe describe in a sentence. I could be mis-remembering, but isn’t each site made up of 50 stops, and presumably you aggregated richness across these stops?

Line 105: I assume all these species have been similarly ID’d over the course of the dataset – e.g. birds split into sub-species haven’t been included.

"Line 107: For the environmental variables, do you have any that match up with the temporal resolution of the response data (monthly?). It seems like all the variables are annual / quarterly / etc."

--See comment above on using shorter time scale environmental variables.

### 2017-11-08 manuscript/reviewer_reply.Rmd

Line 62: I really like the recommendations / clarity in Box 1, and think these will be accessible to most ecologists

"Line 71-73: I’m re-reading this after commenting on data, and am now a little confused. The bird responses are annual snapshots during the breeding season? And I thought the environmental data you used was annual or quarterly, not monthly?
https://www.mbr-pwrc.usgs.gov/bbs/grass/genintro.htm"

--BBS data are generally viewed as annual data, even though the observations happen on a single day. The presence of an individual species is primarily a function of suitable habitat and, for many species, timing of migration. Environmental variables with a shorter temporal scale could likely improve models, such as incorporating cool weather days into the observation model, and would be worth pursuing in a different study.

Methods:
"Line 92: For those not familiar with the BBS protocol, maybe describe in a sentence. I could be mis-remembering, but isn’t each site made up of 50 stops, and presumably you aggregated richness across these stops?"

--This is correct, we've added clarification to the Methods.

"Line 105: I assume all these species have been similarly ID’d over the course of the dataset – e.g. birds split into sub-species haven’t been included."

--This is correct, see the section "Richness Data" in the Methods.

@@

"Line 107: For the environmental variables, do you have any that match up with the temporal resolution of the response data (monthly?). It seems like all the variables are annual / quarterly / etc."

--See comment above on using shorter time scale environmental variables.

Line 124: I’m not as familiar with terrestrial projections, but are the CMIP5 projections relatively unbiased / precise when compared to recent data?

"Line 138: Maybe also cite any of the interfaces you used, like RStan:
http://mc-stan.org/users/citations/"

--RStan is cited in the last paragraph of the Methods.

### 2017-11-08 manuscript/draft.Rmd
clarify bbs methods

We evaluated 6 types of forecasting models (Table 1) by dividing the 32 years
of data into 22 years of training data and 10 years of data for evaluating
forecasts using hindcasting. Here we use definitions from meteorology, where a
hindcast is generally any prediction for an event that has already happened,
while forecasts are predictions for actual future events [@jolliffe2003]. We
also made long term forecasts by using the full data set for training and
making forecasts through the year 2050. For both time frames, we made forecasts
using each model with and without correcting for observer effects, as described
below.

@@

**Richness data.** Bird species richness was obtained from the North American
Breeding Bird Survey (BBS) [@pardieck2017] using the Data Retriever Python
package [@morris2013] and rdataretriever R package [@rdataretriever]. The BBS
data was filtered to exclude all nocturnal, cepuscular, and aquatic species
[since these species are not well sampled by BBS methods; @hurlbert2005], as
well as unidentified species, and hybrids. All data from surveys that did not
meet BBS quality criteria were also excluded.

>>>>

We evaluated 6 types of forecasting models (Table 1) by dividing the 32 years
of data into 22 years of training data and 10 years of data for 
evaluating forecasts using hindcasting. We also made long term forecasts by using
the full data set for training and making forecasts through the year 2050. For
both time scales, we made forecasts using each model with and without
correcting for observer effects, as described below.

@@

**Richness data.** Bird species richness was obtained from the North American
Breeding Bird Survey (BBS) [@pardieck2017] using the Data Retriever Python
package [@morris2013] and rdataretriever R package [@rdataretriever]. BBS
observations are three minute point counts made at 50 fixed locations along a
40km route. Here we denote each route as a site and summarize richness as the
total species observed at all 50 locations in each surveyed year. Prior to
summarizing the data was filtered to exclude all nocturnal, cepuscular, and
aquatic species [since these species are not well sampled by BBS methods;
@hurlbert2005], as well as unidentified species, and hybrids. All data from
surveys that did not meet BBS quality criteria were also excluded.
